---
title: "Private and Public Colleges in the US"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Loading Data

```{r}
library(tidyverse)
```

```{r}
colleges <- read.csv("College_Data.csv")
```

### Wrangling for easier manipulation and visualization

```{r cars}
colleges_percent <- colleges %>%
  mutate(percent_acc = Accept/Apps)
```

```{r}
cost <- colleges_percent %>%
  mutate(total = Room.Board + Books + Personal) %>%
  filter(!Grad.Rate >= 101) %>%
  mutate(re_private = ifelse(Private == "Yes", "Private", "Public")) %>%
  select(-Private) %>%
  rename(Private = re_private)
```

### Number of Private and Public Colleges in the US

```{r}
total_sum <- cost %>%
  group_by(Private) %>%
  summarise(total = n())
knitr::kable(total_sum)
```

## Figure 1: Graduation Rates for private and public colleges
```{r}
ggplot(cost, aes(x = Grad.Rate)) +
  geom_histogram(fill = "dodgerblue4", bins = 15) +
  facet_wrap(~Private, nrow = 2) +
  labs(x = "Graduation Rate")
```

## Summary Statistics for graduation rates
```{r}
grad_rate_sum <- cost %>%
  group_by(Private) %>%
  summarise(median = median(Grad.Rate),
            min = min(Grad.Rate),
            max = max(Grad.Rate))
knitr::kable(grad_rate_sum)
```

  Private colleges appear to have higher graduation rates than public colleges. However, the majority of both types of colleges have graduation rates higher than 35-40%.

## Figure 2: Room and board costs for Private and Public Colleges
```{r}
ggplot(cost, aes(x = Room.Board)) +
  geom_histogram(fill = "dodgerblue4", bins = 15) +
  facet_wrap(~Private, nrow = 2) +
  labs(x = "Room and Board Cost")
```

## Summary Statistics for Room and Board Costs
```{r}
rb_cost_sum <- cost %>%
  group_by(Private) %>%
  summarise(median = median(Room.Board),
            min = min(Room.Board),
            max = max(Room.Board))
knitr::kable(rb_cost_sum)
```

  Private colleges typically charge more for room and board. However, there is a large amount of overlap between private and public colleges around $3000 to $5000.  

Data Source: https://www.kaggle.com/yashgpt/us-college-data
